---
title: "SNP Clustering Step 1: Generate SNP Distance Matrices"
author: "Kyle Gontjes"
date: "`r Sys.Date()`"
output: html_document
---

Notes:

1. This is an old variant calling run

2. Reference genome was a concatenated short read assembly. Although this worked well, it was better to use a closed genome as the reference

3. For transparency, we include this old code, but used the new public genome reference-based variant calling approach in this updated manuscript.

```{r setup, include=FALSE}
require(knitr)
opts_knit$set(echo = T,warning = T, message = T,error=T,comment=NA) 
opts_chunk$set(echo = T,warning = T, message = T,error=T,comment=NA)
opts_knit$set(root.dir='~/Desktop/gl_nfs/Project_MERLIN/')
```
 
# Environment

```{r environment, echo=T, message=FALSE, include=T, results=T}
#Packages
packages <- c("tidyverse",'ape','kableExtra','phytools','igraph','ggraph','tableone','kableExtra') 

#Load Packages
invisible(lapply(packages,library,character.only=T,quietly=T))

# Source functions
source("./Analysis/Household_transmission/MERLIN-prolonged-colonization-ms/lib/common_functions.R")

# Print environment
set.seed(45) 
Sys.info()
sessionInfo()
```

# Load dataset
```{r} 
df <- readRDS("./Analysis/Household_transmission/MERLIN-prolonged-colonization-ms/data/dataset/metadata_w_genomic_id_w_duplicate_ESCrE_addressed_w_qcd_w_sra_metadata.RDS") %>% subset(WGS_escre==T & qc_pass==T)

paste0("Number of considered isolates: ",nrow(df))
``` 

# Load fasta files and create pairwise SNP matrices

```{r}
# Load
setwd("./Sequence_data/variant_calling/2025-04-15_SNPKIT/")

# Get species list
species <- table(df$Species) %>% subset(.>1) %>% names %>% str_split(" ",simplify = T) %>% apply(.,1,FUN=function(x){
  genus <- x[1] %>% substr(.,1,1)  %>% tolower()
  species <- x[2]
  paste0(genus,species)
})  

# Create distance matrices
distance_matrices <- pbmcapply::pbmclapply(species,FUN=function(x){  
  # Load file
  if(x %in% c('cportucalensis','sliquefaciens')){
    name <- paste0(x,'_var_sites.fa')
    path <- paste0("./",x,"/snpkit_rfmt_results/phylo_analysis/IQtree/",name)
  } else {
  name <- paste0(x,"_gubbins_masked_var_sites.fa")
  path <- paste0("./",x,"/snpkit_rfmt_results/phylo_analysis/gubbins_masked/",name)
  }
  # Load DNA
  dna <- read.dna(path,format = 'fasta')
  # Create DNA matrix using ape
  dna_matrix <- ape::dist.dna(x = dna, # DNAbin object as read in above
                       as.matrix = TRUE, # return as matrix
                       model = "N", # count pairwise distances
                       pairwise.deletion = TRUE # delete sites with missing data in a pairwise way
) %>%  as.data.frame 
  if(x == "ecoli"){
    dna_matrix <- dna_matrix %>% select(-'NZ_UFZF01000006.1') %>% subset(rownames(.) != 'NZ_UFZF01000006.1')
  }   
  # Remove unnecessary files 
  keep_ids <- intersect(rownames(dna_matrix), df$genomic_id)
  dna_matrix <- dna_matrix[keep_ids, keep_ids] 
  return(dna_matrix)
},mc.cores = 6) %>% `names<-`(species)
```

# Save file

```{r} 
saveRDS(distance_matrices,'./Analysis/Household_transmission/MERLIN-prolonged-colonization-ms/data/SNP_distances/deprecated/SNP_distance_matrices/MERLIN_SNP_distance_matrices_concatenated_references.RDS')
```
