---
title: "Data Cleaning Step 7 - Curate final metadata files for MERLIN projects"
author: "Kyle Gontjes"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
require(knitr)
opts_knit$set(echo = T,warning = FALSE, message = FALSE,error=F,comment=NA) 
opts_chunk$set(echo = T,warning = FALSE, message = FALSE,error=F,comment=NA)
opts_knit$set(root.dir='~/Desktop/gl_nfs/Project_MERLIN/Analysis/Household_transmission/MERLIN-prolonged-colonization-ms/')
```
 
NOTE: Run the following analysis scripts before doing this!

1. SNP distance analysis: ./scripts/SNP_distance

2. E. coli phylogroup analyses: ./scripts/ecoli_phylogroup


# Environment

```{r environment, echo=T, message=FALSE, include=T, results=T}
# Packages
packages <- c("tidyverse",'kableExtra') 

# Load Packages
invisible(lapply(packages,library,character.only=T,quietly=T))

# Source functions
source("./lib/common_functions.R")

# Print environment
set.seed(45) 
Sys.info()
sessionInfo()
```

# Load data

```{r}  
df <- readRDS("./data/dataset/metadata_w_genomic_id_w_duplicate_ESCrE_addressed_w_qcd_w_sra_metadata.RDS") 
cluster_assignments <- readRDS("./data/SNP_distances/SNP_cluster_assignments/MERLIN_SNP_cluster_assignments_public_references.RDS")
ecoli_phylogroup <- readRDS("./data/ecoli_phylogroup/ecoli_phylogroup_curated.RDS") 
``` 

# Merge clustering data 
## Verify similarity across SNP thresholds from 18 to 31
```{r}
thresholds <- 0:100
SNP_clustering_across_relevant_thresholds <- cluster_assignments %>% select(paste0("Species_ST_cluster_",thresholds))

lapply(thresholds,FUN=function(x){
  cluster_summary_statistics(SNP_clustering_across_relevant_thresholds[[paste0("Species_ST_cluster_",x)]]) %>% mutate(threshold=x) %>% select(threshold,everything())
}) %>% do.call(rbind,.) %>% kable() %>% kable_classic_2()
```


## Cluster assignments use a 18 SNP threshold (single linkage); renamed to Species_ST_cluster for consistency with prior scripts

```{r}
df <- left_join(df, cluster_assignments %>% 
                  select(genomic_id,Species_ST_cluster_18) %>%
                  rename(Species_ST_cluster = Species_ST_cluster_18))
```

# Merge E. coli Phylogroup data
## For genomes with E. coli, they should provide an E. coli phylogroup
## If genome is not Escherichia, this would be clarified as "Not Escherichia"

```{r}
# Function
get_clermont_data <- function(genomic_id, clermont_data){
  ids <- str_split(genomic_id,";") %>% unlist
  clermont_sub <- clermont_data %>% subset(genomic_id %in% ids)
  if(nrow(clermont_sub) >0 ){
    clermont_sub[["ecoli_phylogroup"]] %>% paste0(collapse=";")
    } else{
    NA
    }
}

# Implementation
df$ecoli_phylogroup <- sapply(df$genomic_id, get_clermont_data, clermont_data = ecoli_phylogroup)
```
 
# Save the dataset!

```{r}
saveRDS(df,"./data/dataset/MERLIN_cohort_metadata_final.RDS")
write_csv(df,"./data/dataset/MERLIN_cohort_metadata_final.csv")
```