---
title: "SNP Clustering Step 3: Generate SNP distance clustering across thresholds for concatenated reference genomes"
author: "Kyle Gontjes"
date: "`r Sys.Date()`"
output: html_document
---

Notes:

1. This is an old variant calling run

2. Reference genome was a concatenated short read assembly. Although this worked well, it was better to use a closed genome as the reference

3. For transparency, we include this old code, but used the new public genome reference-based variant calling approach in this updated manuscript.

```{r setup, include=FALSE}
require(knitr)
opts_knit$set(echo = T,warning = FALSE, message = FALSE,error=F,comment=NA) 
opts_chunk$set(echo = T,warning = FALSE, message = FALSE,error=F,comment=NA)
opts_knit$set(root.dir='~/Desktop/gl_nfs/Project_MERLIN/Analysis/Household_transmission/MERLIN-prolonged-colonization-ms/')
```
 
# Environment

```{r environment, echo=T, message=FALSE, include=T, results=T}
#Packages
packages <- c("tidyverse",'ggraph','igraph','kableExtra') 

#Load Packages
invisible(lapply(packages,library,character.only=T,quietly=T))

#Source functions
source("./lib/common_functions.R")

# Print environment
set.seed(45) 
Sys.info()
sessionInfo()
```

# Load data

```{r}  
df <- readRDS("./data/dataset/metadata_w_genomic_id_w_duplicate_ESCrE_addressed_w_qcd_w_sra_metadata.RDS") %>% subset(WGS_escre==T & qc_pass==T)

SNP_distance_pair_dataset_concatenated_references <- readRDS('./data/SNP_distances/deprecated/SNP_pairwise_datasets/MERLIN_SNP_distance_pair_dataset_concatenated_references.RDS') 
``` 

# Generate clustering data 

```{r}
thresholds_of_interest <- c(0:100) 
ST_list <- df$Species_ST %>% unique %>% sort

cluster_calls_concatenated_reference <- pbmcapply::pbmclapply(thresholds_of_interest,FUN=function(x){
  # Get SNP threhsold clustering
  threshold_data <- SNP_threshold_based_single_linkage_clustering(dataframe = df,pair_types =  SNP_distance_pair_dataset_concatenated_references, threshold = x, distance_variable = 'pairwise_dist') 
  # Sequence type clusteirng
  threshold_data_w_annotations <- left_join(threshold_data,df %>% select(genomic_id,Species,Species_ST,ST))
  ST_df <- lapply(ST_list,recode_clustering_calls_using_ST_data,threshold_data_w_annotations) %>% do.call(rbind,.)  %>%
  rename(
    !!paste0("cluster_", x) := cluster,
    !!paste0("Species_ST_cluster_", x) := Species_ST_cluster
  )
  ST_df
}, mc.cores=10)  
``` 

# Save cluster calls across thresholds

```{r}
cluster_calls_across_thresholds <-  lapply(cluster_calls_concatenated_reference,FUN=function(x){x %>% select(-clustering_distance_variable)}) %>%
  reduce(left_join, by=c("genomic_id", "Species", "Species_ST", "ST")) %>%
  select(genomic_id, Species, ST, Species_ST, 
         sort(starts_with("cluster_")), 
         sort(starts_with("Species_ST_cluster_"))) 

saveRDS(cluster_calls_across_thresholds,'./data/SNP_distances/deprecated/SNP_cluster_assignments/MERLIN_SNP_cluster_assignments_concatenated_references.RDS')
write_csv(cluster_calls_across_thresholds,'./data/SNP_distances/deprecated/SNP_cluster_assignments/MERLIN_SNP_cluster_assignments_concatenated_references.csv') 
```